---
title: "HW2"
author: "Yiting Zhang"
date: '2022-04-10'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```



## Linear Regression

For this lab, we will be working with a data set from the UCI (University of California, Irvine) Machine Learning repository ([see website here](http://archive.ics.uci.edu/ml/datasets/Abalone)). The full data set consists of $4,177$ observations of abalone in Tasmania. (Fun fact: [Tasmania](https://en.wikipedia.org/wiki/Tasmania "Tasmania") supplies about $25\%$ of the yearly world abalone harvest.)


The age of an abalone is typically determined by cutting the shell open and counting the number of rings with a microscope. The purpose of this data set is to determine whether abalone age (**number of rings + 1.5**) can be accurately predicted using other, easier-to-obtain information about the abalone.

The full abalone data set is located in the `\data` subdirectory. Read it into *R* using `read_csv()`. Take a moment to read through the codebook (`abalone_codebook.txt`) and familiarize yourself with the variable definitions.

Make sure you load the `tidyverse` and `tidymodels`!


```{r}
library(tidyverse)
library(dplyr)
library(tidymodels)
library(ISLR)
library(yardstick)
tidymodels_prefer()
```

```{r}
abalone <- read.csv("abalone.csv")
```

### Question 1

Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no `age` variable in the data set. Add `age` to the data set.

Assess and describe the distribution of `age`.

```{r}
abalone <- abalone%>%
  mutate(age = abalone$rings + 1.5)

ggplot(data=abalone, aes(age)) + geom_histogram()
```
By plotting a histogram, it seems like the abalone age data follows a normal distribution which skewed to the right a little bit with center around 10.


### Question 2

Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.

*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*

```{r}
set.seed(100)
abalone_split <- initial_split(abalone, prop = 0.80, strata = age)
abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)
```

### Question 3

Using the **training** data, create a recipe predicting the outcome variable, `age`, with all other predictor variables. Note that you should not include `rings` to predict `age`. Explain why you shouldn't use `rings` to predict `age`.

Steps for your recipe:

1.  dummy code any categorical predictors

2.  create interactions between

    -   `type` and `shucked_weight`,
    -   `longest_shell` and `diameter`,
    -   `shucked_weight` and `shell_weight`

3.  center all predictors, and

4.  scale all predictors.

You'll need to investigate the `tidymodels` documentation to find the appropriate step functions to use.

```{r}
abalone_recipe <- recipe(age ~ type + longest_shell + diameter + height + 
                           whole_weight + shucked_weight +viscera_weight + 
                           shell_weight, data = abalone_train) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_interact(terms = ~ type:shucked_weight) %>% 
  step_interact(terms = ~ longest_shell:diameter) %>% 
  step_interact(terms = ~ shucked_weight:shell_weight) %>% 
  step_center(all_predictors()) %>% 
  step_normalize() %>%
  step_scale()
abalone_recipe
```

Rings cannot be used to predict age because it is used in the formula for the age variable. If it is used, then the prediction will be 100% correct.


### Question 4

Create and store a linear regression object using the `"lm"` engine.

```{r}
lm_model <- linear_reg() %>% 
  set_engine("lm")
```

### Question 5

Now:

1.  set up an empty workflow,
2.  add the model you created in Question 4, and
3.  add the recipe that you created in Question 3.

```{r}
lm_wflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(abalone_recipe)
```

### Question 6

Use your `fit()` object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.


```{r}
lm_fit <- lm_wflow %>% fit(abalone_train)

Q6_data <- data.frame(type="F",longest_shell = 0.50, diameter = 0.10,
                      height = 0.30, whole_weight = 4, shucked_weight = 1, 
                      viscera_weight = 2, shell_weight = 1,stringsAsFactors = TRUE)


prediction <-predict(lm_fit, new_data = Q6_data)

prediction
```


### Question 7

Now you want to assess your model's performance. To do this, use the `yardstick` package:

1.  Create a metric set that includes *R^2^*, RMSE (root mean squared error), and MAE (mean absolute error).
2.  Use `predict()` and `bind_cols()` to create a tibble of your model's predicted values from the **training data** along with the actual observed ages (these are needed to assess your model's performance).
3.  Finally, apply your metric set to the tibble, report the results, and interpret the *R^2^* value.

```{r}

abalone_train_res <- bind_cols(predict(lm_fit, abalone_train), abalone_train$age)
  
abalone_metrics <- metric_set(rsq, rmse, mae)
abalone_metrics(abalone_train_res, truth = ...2, 
                estimate = .pred)
```

The $R^2$ value is 0.5498393, which means that the linear regression model we made explains 54.98393% of the variation in age variable.